paper

### 2024-09-16 paper.qmd
.

## Run t-test

```{r}
library(SummarizedExperiment, quietly = T, warn.conflicts = F)
data("UPS1.Case0")
data <- assay(UPS1.Case0)
meta.info <- data.frame(colData(UPS1.Case0), check.names = FALSE, row.names = row.names(colData(UPS1.Case0)))
t.test.result <- data.frame()

for(i in 1:nrow(data)){
  p <- t.test(data[i,1:12], data[i,13:24])
  t.test.result <- rbind(t.test.result, data.frame(protein = row.names(data)[i] ,p.value = p$p.value))
}

t.test.result$fdr <- p.adjust(t.test.result$p.value)
saveRDS(t.test.result, "../LimROTS paper/UPS1.Case0.t.test.rds")
```

@@

limrots.result <- readRDS("../LimROTS paper/UPS1.Case0.LimROTS.rds")
rots.result <- readRDS("../LimROTS paper/UPS1.Case0.ROTS.rds")
limma.result <- readRDS("../LimROTS paper/UPS1.Case0.limma.rds")
t.test.result <- readRDS("../LimROTS paper/UPS1.Case0.t.test.rds")

@@

predictions_t.test <- t.test.result$fdr < 0.01
predictions_t.test <- factor(predictions_t.test, levels = c(TRUE, FALSE))
true_labels_t.test <- ifelse(t.test.result$protein %in% TP , TRUE, FALSE)
true_labels_t.test <- factor(true_labels_t.test, levels = c(TRUE, FALSE))
conf_matrix_t.test <- confusionMatrix(predictions_t.test, true_labels_t.test)

@@

```{r}
library(dplyr, quietly = T, warn.conflicts = T)  
library(ggplot2)
library(ggsci)

@@

metrics_limrots <- extract_metrics(conf_matrix_limrots, "limROTS")
metrics_rots <- extract_metrics(conf_matrix_rots, "ROTS")
metrics_limma <- extract_metrics(conf_matrix_limma, "limma")
metrics_t.test <- extract_metrics(conf_matrix_t.test, "t.test")

all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma,metrics_t.test)

ggplot(all_metrics, aes(x = Metric, y = Value, fill = Method)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  labs(title = "Comparison of Performance Case Study0",
       y = "Value", x = "Metric") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_npg() + 
    theme(
    plot.title = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.text = element_text(face = "bold", color = "black"),
    axis.ticks = element_line(color = "black")
  )

@@

>>>>

@@

@@

@@

@@

metrics_limrots <- extract_metrics(conf_matrix_limrots, "limROTS")
metrics_rots <- extract_metrics(conf_matrix_rots, "ROTS")
metrics_limma <- extract_metrics(conf_matrix_limma, "limma")

all_metrics <- bind_rows(metrics_limrots, metrics_rots, metrics_limma)

ggplot(all_metrics, aes(x = Metric, y = Value, fill = Method)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  labs(title = "Comparison of Performance Metrics",
       y = "Value", x = "Metric") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

# BM21 Case study 3

```{r}
# Load necessary packages
library(LimROTS)
library(parallel)

# Load the dataset
load("../LimROTS paper/metabolomics/BM21.Case0.rda")

print(BM21.Case0)

# Set metadata and formula for LimROTS analysis
meta.info <- "mixing.ratio"
B <- 1000  # Number of bootstrap samples
K <- dim(BM21.Case0)[1]/2  # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 10  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "mixing.ratio"
formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison

# Run LimROTS analysis with trend and robust settings enabled
limrots.result <- LimROTS(data.exp = BM21.Case0, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE)
```

```{r}
length(which(limrots.result$FDR < 0.05))
summary(limrots.result$q_values)
plot(limrots.result$q_values)
hist(limrots.result$FDR)

limrots.result.df <- data.frame(proteins = row.names(limrots.result$data),
                                LimROTS.FC = limrots.result$corrected.logf,
                                FDR = limrots.result$FDR)
```

```{r}
# Load the dataset
load("../LimROTS paper/metabolomics/BM21.Case0.rda")
  library(SummarizedExperiment)
library(doParallel)
library(LimROTS)
print(BM21.Case0)

# Set metadata and formula for LimROTS analysis
meta.info <- "mixing.ratio"
B <- 1000  # Number of bootstrap samples
K <- dim(BM21.Case0)[1]/2 # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 10  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "mixing.ratio"
formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison

# Run LimROTS analysis with trend and robust settings enabled
rots.result <- LimROTS(data.exp = BM21.Case0, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = T)
```

```{r}
length(which(rots.result$FDR < 0.01))
summary(rots.result$q_values)
plot(rots.result$q_values)
hist(rots.result$q_values)
hist(rots.result$p)
hist(rots.result$FDR)
```

## Run Limma

```{r}

library(limma)
library(SummarizedExperiment, quietly = T, warn.conflicts = T)

load("../LimROTS paper/metabolomics/BM21.Case0.rda")

print("BM21.Case0")
data <- assay(BM21.Case0)
formula.str <- "~ 0 + mixing.ratio"
meta.info <- data.frame(colData(BM21.Case0), check.names = FALSE, row.names = row.names(colData(BM21.Case0)))
meta.info$sample.id <- NULL
design.matrix <- model.matrix(formula(formula.str), data = meta.info)
colnames(design.matrix) <- make.names(colnames(design.matrix))
fit <- lmFit(data, design.matrix)
cont_matrix <- makeContrasts("mixing.ratio1-mixing.ratio2",  levels=design.matrix)
fit2 <- contrasts.fit(fit, cont_matrix)
fit.ebayes <- eBayes(fit2, trend=T, robust=T)
limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")

length(which(limma.result$adj.P.Val < 0.01))

saveRDS(limma.result, "../LimROTS paper/BM21.case0.limma.rds")
```
